Reconstructing the last known movements of one of Nature's giants

### 2017-06-30 manuscript/blue-whale-ms.tex
Change data to zenodo link
Final edits NC

\paragraph{Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change\cite{runge2014conserving,robinson2009travelling}. 
Animal tissues housed in museum collections provide biochemical records of behaviour that can be used to reconstruct individual-level movements over long timescales and under historical climatic and ecological conditions\cite{newsome2010using}. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to reconstruct the movements of an iconic blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London (NHM). 
We identify two distinct phases in the whale's behaviour: repeated annual north-south migrations in sub-Arctic waters, and residency in subtropical waters. 
Subtropical residency is associated with isotopic signatures consistent with pregnancy and lactation in the year before the whale's death. 
These results supply the first quantitative multi-year records of historical migrations in individual blue whales, and suggest that blue whale residency and movement patterns in the North Atlantic have remained similar for hundreds of years, despite changing human pressures in the region. 
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climate, ecological conditions and anthropogenic pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modelling unlocks the information recorded in these irreplaceable archives.}

@@

The blue whale (\textit{Balaenoptera musculus}), the largest animal to have ever lived, is one such migratory marine species. 
Despite blue whale hunting being banned for over 50 years, blue whales are still listed as Endangered by the IUCN Red List\cite{reilly2008balaenoptera}.
Conservation efforts are hampered by the scarcity of detailed data on the distribution, behaviour and ecology of blue whales, both for historical and present-day populations. 
Current methods for tracking large whales include satellite tracking, radio tagging, aerial photography, passive acoustic monitoring, and sightings records\cite{borger15,mcdonald2006biogeographic,bailey2009behavioural,mate2007evolution,silva2013north}. 
These methods are adding vital behavioural data to our understanding of whale movements, however, they are often expensive, limited to short-term studies on few individuals\cite{bailey2009behavioural,best2015tag,mate2007evolution}, and may influence whale behaviour\cite{walker2012review}. 
Intrinsic biogeochemical tags provide an alternative technique, where the chemical compositions of animal tissues are compared to spatially referenced models of chemical variation across the species' range\cite{hobson1999tracing,hobson2008tracking}). 
Stable isotope-based geolocation has been used to interpret whale ecology \cite{hobson1998stable,ryan2013stable,eisenmann2016isotopic}. 
Here we apply this method to baleen from the Natural History Museum's blue whale.

@@

% figure 1
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-1-raw-dC-dN-data.png}
  \caption{Variation in $\delta^{13}$C (black circles, left y-axis) and $\delta^{15}$N (grey circles, right x-axis) longitudinally through the baleen plate (n = 77 samples from a single baleen plate for both isotopes) of the NHM blue whale exhibits strong annual periodicity and cross-correlation (Figs. S1 and S2) in both isotopes as well as a change in the mean and magnitude of the cycles during 1890 (approximately at sample 30cm). Approximate relationship to years assuming a growth rate of 13.5 cm y$^{-1}$ is shown on the upper x-axis, and year boundaries are indicated by vertical dotted grey lines.}
  \label{fig1}
\end{figure}

Relating marine animal tissue isotopic compositions to likely locations is complicated by both a lack of spatial reference values, and because the isotopic composition of phytoplankton at the base of marine food webs varies in both space and time. 
We approach these problems using newly developed, temporally dynamic models of $\delta^{13}$C values in the global ocean\cite{magozzi2017using} (Fig. S3), coupled to an agent-based model of whale movements (see Supplementary Methods; Table S1) to simulate time series of $\delta^{13}$C values expected under differing movement behaviours and in differing parts of the known geographic range of blue whales. 
We then compare simulated profiles (Figs. \ref{fig2}, \ref{fig3} and Fig. S4) with the measured isotopic profile of the blue whale's baleen (Fig. \ref{fig1}) to identify the most likely movement history.

% figure 2
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-2-migratory-model-full-map.png}
  \caption{Simulated tracks from the migratory movement model (2250 days per each of 50 replicates) with simulated day indicated by colour illustrating the residency period from Ireland to the Norwegian/Barents Sea followed by a migratory period and year spent near the Azores (as indicated by an arrow) and/or Canary Islands before returning north prior to stranding in Wexford, Ireland 1891 (as indicated by an arrow). }
  \label{fig2}
\end{figure}

% figure 3
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-3-migratory-model-d13C.png}
  \caption{The corresponding $\delta^{13}$C data from these simulated movements correlate with the empirical data from the Baleen: across the 50 simulations, the correlation coefficient between z-score transformations of empirical and time-matched simulation values had a mean of 0.48 ($sd = 0.2$), 49 of them were positive,  and 39 of the 50 Pearson correlation tests had a p-value less than the Bonferroni corrected $p = 0.05/50 = 0.001$. The initiation of the southward migration in the late spring of 1889 is indicated by the vertical broken line. The standard deviation of the empirical data have been rescaled by a factor of three so as to better visualise the correlation since as a consumer, the whale is expected to integrate and attenuate the fluctuations in its diet, and the end points of the simulations and empirical data have been aligned to coincide.}
  \label{fig3}
\end{figure}

@@

Conventional wisdom suggests that baleen whales migrate each year from feeding grounds near the poles in summer, to overwintering and breeding grounds nearer the equator\cite{corkeron1999baleen,lockyer1981migration}. 
However, a growing body of evidence suggests that not all whales migrate to subtropical waters each year, with some remaining at high latitudes year round\cite{mcdonald2006biogeographic}. 
Our results support this conclusion. 
Interestingly, there is still a resident population of blue whales in the regions indicated by our model, and females still breed in the subtropical waters near the Azores\cite{reilly2008balaenoptera}. 
Whaling was an intense pressure for blue whales during the period we are analysing. 
Before whaling in the North Atlantic began in 1868\cite{reilly2008balaenoptera}, there were 10,000-15,000 blue whales in the region\cite{sigurjonsson1995life}. 
In the early 20th century, fisheries moved outside the area because stocks were so depleted\cite{reilly2008balaenoptera}; during this period over 12,000 blue whales were landed\cite{sigurjonsson1995life}. 
The NHM blue whale thus represents an individual from a species at the brink of extinction.
Our data suggest that the spatial ecology of blue whales in the North Atlantic have remained similar for more than 100 years, despite dramatic changes in anthropogenic pressures.

This is the first quantitative evidence of historical individual-level movement behaviours in blue whales. 
Our novel simulation modelling removes a long-standing limitation in stable isotope ecology, and can be applied to stable isotope records from any incrementally-grown tissue to estimate most likely individual movement behaviours over multiple years. 
Such techniques can also be applied to extinct species' behaviours and used to understand how populations were affected by past pressures such as hunting, and present-day pressures such as global change. 
Finally, our results highlight that museums represent a vast, mostly untapped, source of individual-level spatial ecology information.

@@

\subsection{Stable isotope extractions from baleen}
Baleen was collected from the Natural History Museum, London (specimen NHMUK.1892.3.1.1). 
The baleen plate was cleaned with ethanol to remove surface contaminants. 
{\raise.17ex\hbox{$\scriptstyle\sim$}}1mg samples of keratin powder were then collected from the plate using a hand-held drill and grinding bit. 
78 samples were taken at 1cm intervals, 0.5cm from the outer edge of the plate, starting at the proximal section which contains the most recent tissue. 
Baleen grows at a constant rate, so the samples are equally spaced through time\cite{best1996stable}. Carbon and nitrogen isotope analysis was performed simultaneously via continuous-flow isotope ratio mass spectrometry at the National Oceanographic Centre (Southampton, UK), using a Vario Isotope select elemental analyser, coupled with Isoprime 100 isotope mass spectrometer. 
Replicates using internal laboratory standards (L-glutamic acid (C), Glutamic acid (CT standard), acetanilide and protein standard OAS) were used for quality control and calibration. 
C:N ratios for samples ranged from 3.28\text{\textperthousand} to 3.72\text{\textperthousand}, well within the acceptable theoretical range for pure keratin ($3.4\pm0.5$) allowing for comparison among samples (Hobson  Koch 1998).

\subsection{Time calibrating stable isotope profiles}
Seasonal migrations across isotopic gradients induce cyclical variations in the isotopic composition of baleen, the distance between cycles reflecting growth rates\cite{hobson1998stable,busquets2017estimating}. 
Clear periodicity was evident in $\delta^{15}$N values across the entire baleen plate, and in $\delta^{13}$C values in the most distal (oldest) 65cm of the plate. 
We calculated isotopic periodicity within the baleen sample using Fourier Transform analysis\cite{cardona2017temporal} (Fig. S?), revealing a consistent growth rate of 13.5cm $y^{-1}$ which is remarkably similar to the mean isotope-derived baleen growth rates of $15.5 \pm 2.2cm y{^-1}$ estimated for six recent blue whales from the Californian Pacific\cite{busquets2017estimating}.  
Therefore we dated the youngest baleen sample as 1st March 1891, 24 days prior to the stranding date, 25th March 1891. 
Cross-correlation analysis demonstrated a strong negative covariance between $\delta^{13}$C and $\delta^{15}$N values within behavioural phase one (oldest 60cm of the baleen plate), but no relationship between $\delta^{13}$C and $\delta^{15}$N values during behavioural phase two (Fig. S?).

@@

\subsection{Agent-based whale movement model}
We simulated the movement of the whale with the likelihood, direction, and extent of movement influenced by behavioural state, sea surface temperature, water depth, and phytoplankton concentration (as a proxy for zooplankton food availability). 
At each time point and location in the simulations, we extracted phytoplankton $\delta^{13}$C values from the spatio-temporal model\cite{magozzi2017using} resulting in a range of simulated isotopic profiles for different migration patterns and foraging areas. 
Movement was coded as a set of probabilistic rules. 
The parameters for these were taken from the literature on blue whale behaviour (e.g.\cite{handbook}). 
All terms were expressed as probability distributions allowing us to introduce individual variability.

In the models, the likelihood of movement, direction (north, south, east, west, northeast, northwest, southeast, or southwest) and extent (1$^{\circ}$ or 2$^{\circ}$ in km) of movement are all influenced by the following. 
(i) Behavioural state (migrating north, migrating south, foraging, or nursing). 
Northerly migrations were possible only in spring, and southerly migrations in autumn. 
Foraging was possible at any time of year, and was triggered when whales encountered high concentrations of plankton. Nursing occurred at temperatures between 18 and 25$^{\circ}$C only. 
(ii) Sea surface temperature ($^{\circ}$C\cite{yool2013medusa}). 
When migrating north, whales were more likely to move towards lower temperatures provided they were above the minimum temperature threshold (3$^{\circ}$C); whereas whales migrating south sought warmer waters. 
(iii) Water depth (m; from the General Bathymetric Chart of the Oceans (GEBCO) global bathymetry dataset). 
Whales were less likely to move into waters less than 400m deep, and increasingly unlikely to move into even shallower waters. 
(iv) Phytoplankton concentration (mmol N $m^{-3}$, for combined diatom and non-diatom communities\cite{yool2013medusa}). 
This was included as a proxy for zooplankton food availability. 
Whales are more likely to move towards (or remain within) areas of high phytoplankton density, particularly during the foraging behavioural state.

We simulated the isotopic expression expected for (a) residency in each known hotspot for blue whale sightings or historic hunting grounds in the North Atlantic (Norwegian/Barents Sea, West Ireland, South Greenland, Nova Scotia/Newfoundland and Canaries/Azores\cite{mcdonald2006biogeographic,reilly2008balaenoptera,sigurjonsson1995life}; Fig S4); (b) seasonal migration between high sub-arctic latitudes and temperate latitudes around the British Isles and (c) seasonal migration between high latitudes and subtropical latitudes. 
For a full description of the model and its parameters see Supplementary Methods and Table S1. 
We simulated whale movements 30 times for each scenario and residency hotspot (Fig. S4). 
For each movement simulation we extracted phytoplankton $\delta^{13}$C values at each time point and location from the models described above\cite{magozzi2017using}. 
We then compared the simulated stable isotope profiles (Fig. \ref{fig3}) to the profile of the blue whale (Fig. \ref{fig1}).

\subsection{Data Availability}
Data are available at XXX. 
R code is available from GitHub (github.com/ALJackson/blue-whale-bes).

>>>>

\paragraph{Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change\cite{runge2014conserving,robinson2009travelling}. 
Animal tissues housed in museum collections provide biochemical records of behaviour that can be used to reconstruct individual-level movements over long timescales and under historical climatic and ecological conditions\cite{newsome2010using}. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to reconstruct the movements of an iconic blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London (NHM). 
We identify two distinct phases in the whale's behaviour: repeated annual north-south migrations in sub-Arctic waters, and residency in subtropical waters. 
Subtropical residency is associated with isotopic signatures consistent with pregnancy and lactation in the year before the whale's death. 
These results supply the first quantitative multi-year records of historical migrations in individual blue whales, and suggest that blue whale residency and movement patterns in the North Atlantic have remained similar for hundreds of years, despite changing human pressures in the region. 
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climate, ecological conditions and anthropogenic pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modeling unlocks the information recorded in these irreplaceable archives.}

@@

The blue whale (\textit{Balaenoptera musculus}), the largest animal to have ever lived, is one such migratory marine species. 
Despite blue whale hunting being banned for over 50 years, blue whales are still listed as Endangered by the IUCN Red List\cite{reilly2008balaenoptera}.
Conservation efforts are hampered by the scarcity of detailed data on the distribution, behaviour and ecology of blue whales, both for historical and present-day populations. 
Current methods for tracking large whales include satellite tracking, radio tagging, aerial photography, passive acoustic monitoring, and sightings records\cite{borger15,mcdonald2006biogeographic,bailey2009behavioural,mate2007evolution,silva2013north}. 
These methods are adding vital behavioural data to our understanding of whale movements, however, they are often expensive, limited to short-term studies on few individuals\cite{bailey2009behavioural,best2015tag,mate2007evolution}, and may influence whale behaviour\cite{walker2012review}. 
Intrinsic biogeochemical tags provide an alternative technique, where the chemical compositions of animal tissues are compared to spatially referenced models of chemical variation across the species' range\cite{hobson1999tracing,hobson2008tracking}). 
Stable isotope-based geolocation has been used to interpret whale ecology \cite{hobson1998stable,ryan2013stable,eisenmann2016isotopic}. 
Here we apply this method to baleen from the NHM's blue whale.

@@

% figure 1
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-1-raw-dC-dN-data.png}
  \caption{Variation in $\delta^{13}$C (black circles, left y-axis) and $\delta^{15}$N (grey circles, right y-axis) longitudinally through the baleen plate (n = 77 samples from a single baleen plate for both isotopes) of the NHM blue whale exhibits strong annual periodicity and cross-correlation (Figs. S1 and S2) in both isotopes as well as a change in the mean and magnitude of the cycles during 1889 (approximately at sample 30cm). Approximate relationship to years assuming a growth rate of 13.5 cm y$^{-1}$ is shown on the upper x-axis, and year boundaries are indicated by vertical dotted grey lines.}
  \label{fig1}
\end{figure}

Relating marine animal tissue isotopic compositions to likely locations is complicated by both a lack of spatial reference values, and because the isotopic composition of phytoplankton at the base of marine food webs varies in both space and time. 
We approach these problems using newly developed, temporally dynamic models of $\delta^{13}$C values in the global ocean\cite{magozzi2017using} (Fig. S3), coupled to an agent-based model of whale movements (see Supplementary Methods; Table S1; Fig. S5) to simulate time series of $\delta^{13}$C values expected under differing movement behaviours and in differing parts of the known geographic range of blue whales. 
We then compare simulated profiles (Figs. \ref{fig2}, \ref{fig3} and Fig. S4) with the measured isotopic profile of the blue whale's baleen (Fig. \ref{fig1}) to identify the most likely movement history.

% figure 2
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-2-migratory-model-full-map.png}
  \caption{Simulated tracks from the migratory movement model (2250 days for each of 50 replicates) with simulated day indicated by colour illustrating the residency period from Ireland to the Norwegian/Barents Sea followed by a migratory period and year spent near the Azores (as indicated by an arrow) and/or Canary Islands before returning north prior to stranding in Wexford, Ireland 1891 (as indicated by an arrow). }
  \label{fig2}
\end{figure}

% figure 3
\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{figures/Figure-3-migratory-model-d13C.png}
  \caption{The corresponding $\delta^{13}$C data from these simulated movements correlate with the empirical data from the baleen: across the 50 simulations, the correlation coefficient between z-score transformations of empirical and time-matched simulation values had a mean of 0.48 ($sd = 0.2$), 49 of them were positive, and 39 of the 50 Pearson correlation tests had a p-value less than the Bonferroni corrected value of 0.001 ($p = 0.05/50 = 0.001$). The initiation of the southward migration in the late spring of 1889 is indicated by the vertical broken line. The standard deviation of the empirical data have been rescaled by a factor of three so as to better visualise the correlation because, as a consumer, the whale is expected to integrate and attenuate the fluctuations in its diet, and the end points of the simulations and empirical data have been aligned to coincide.}
  \label{fig3}
\end{figure}

@@

Conventional wisdom suggests that baleen whales migrate each year from feeding grounds near the poles in summer, to overwintering and breeding grounds nearer the equator\cite{corkeron1999baleen,lockyer1981migration}. 
However, a growing body of evidence suggests that not all whales migrate to subtropical waters each year, with some remaining at high latitudes year round\cite{mcdonald2006biogeographic}. 
Our results support this conclusion. 
Interestingly, there is still a resident population of blue whales in the regions indicated by our model, and females still breed in the subtropical waters near the Azores\cite{reilly2008balaenoptera}. 
Whaling was an intense pressure for blue whales during the period we are analysing. 
Before whaling in the North Atlantic began in 1868\cite{reilly2008balaenoptera}, there were 10,000-15,000 blue whales in the region\cite{sigurjonsson1995life}. 
In the early 20th century, fisheries moved outside the area because stocks were so depleted\cite{reilly2008balaenoptera}; during this period over 12,000 blue whales were landed\cite{sigurjonsson1995life}. 
The NHM blue whale thus represents an individual from a species at the brink of extinction.
Our data suggest that the spatial ecology of blue whales in the North Atlantic has remained similar for more than 100 years, despite dramatic changes in anthropogenic pressures.

This is the first quantitative evidence of historical individual-level movement behaviours in blue whales. 
Our novel simulation modeling removes a long-standing limitation in stable isotope ecology, and can be applied to stable isotope records from any incrementally-grown tissue to estimate most likely individual movement behaviours over multiple years. 
Such techniques can also be applied to extinct species behaviours and used to understand how populations were affected by past pressures such as hunting, and present-day pressures such as global change. 
Finally, our results highlight that museums represent a vast, mostly untapped, source of individual-level spatial ecology information.

@@

\subsection{Stable isotope extractions from baleen}
Baleen was collected from the Natural History Museum, London (specimen NHMUK.1892.3.1.1). 
The baleen plate was cleaned with ethanol to remove surface contaminants. 
{\raise.17ex\hbox{$\scriptstyle\sim$}}1mg samples of keratin powder were then collected from the plate using a hand-held drill and grinding bit. 
77 samples were taken at 1cm intervals, 0.5cm from the outer edge of the plate, starting at the proximal section which contains the most recent tissue. 
Baleen grows at a constant rate, so the samples are equally spaced through time\cite{best1996stable}. Carbon and nitrogen isotope analysis was performed simultaneously via continuous-flow isotope ratio mass spectrometry at the National Oceanographic Centre (Southampton, UK), using a Vario Isotope select elemental analyser, coupled with Isoprime 100 isotope mass spectrometer. 
Replicates using internal laboratory standards (L-glutamic acid (C), Glutamic acid (CT standard), acetanilide and protein standard OAS) were used for quality control and calibration. 
C:N ratios for samples ranged from 3.28\text{\textperthousand} to 3.72\text{\textperthousand}, well within the acceptable theoretical range for pure keratin ($3.4\pm0.5$) allowing for comparison among samples\cite{hobson1998stable}.

\subsection{Time calibrating stable isotope profiles}
Seasonal migrations across isotopic gradients induce cyclical variations in the isotopic composition of baleen, the distance between cycles reflecting growth rates\cite{hobson1998stable,busquets2017estimating}. 
Clear periodicity was evident in $\delta^{15}$N values across the entire baleen plate, and in $\delta^{13}$C values in the most distal (oldest) 65cm of the plate. 
We calculated isotopic periodicity within the baleen sample using Fourier Transform analysis\cite{cardona2017temporal} (Fig. S1), revealing a consistent growth rate of 13.5$cmy^{-1}$ which is remarkably similar to the mean isotope-derived baleen growth rates of $15.5 \pm 2.2cmy^{-1}$ estimated for six recent blue whales from the Californian Pacific\cite{busquets2017estimating}.  
Therefore we dated the youngest baleen sample as 1st March 1891, 24 days prior to the stranding date, 25th March 1891. 
Cross-correlation analysis demonstrated a strong negative covariance between $\delta^{13}$C and $\delta^{15}$N values within behavioural phase one (oldest 60cm of the baleen plate), but no relationship between $\delta^{13}$C and $\delta^{15}$N values during behavioural phase two (Fig. S2).

@@

\subsection{Agent-based whale movement model}
We simulated the movement of the whale with the likelihood, direction, and extent of movement influenced by behavioural state, sea surface temperature, water depth, and phytoplankton concentration (as a proxy for zooplankton food availability). 
At each time point and location in the simulations, we extracted phytoplankton $\delta^{13}$C values from the spatio-temporal model\cite{magozzi2017using} resulting in a range of simulated isotopic profiles for different migration patterns and foraging areas (Figs 2, 3, and S4). 
Movement was coded as a set of probabilistic rules. 
The parameters for these were taken from the literature on blue whale behaviour (e.g.\cite{handbook}). 
All terms were expressed as probability distributions allowing us to introduce individual variability.

In the models, the likelihood of movement, direction (north, south, east, west, northeast, northwest, southeast, or southwest) and extent (1$^{\circ}$ or 2$^{\circ}$ in km) of movement are all influenced by the following. 
(i) Behavioural state (migrating north, migrating south, foraging, or nursing). 
Northerly migrations were possible only in spring, and southerly migrations in autumn. 
Foraging was possible at any time of year, and was triggered when whales encountered high concentrations of plankton. Nursing occurred at temperatures between 18 and 25$^{\circ}$C only. 
(ii) Sea surface temperature\cite{yool2013medusa} ($^{\circ}$C). 
When migrating north, whales were more likely to move towards lower temperatures provided they were above the minimum temperature threshold (3$^{\circ}$C); whereas whales migrating south sought warmer waters. 
(iii) Water depth (m; from the General Bathymetric Chart of the Oceans (GEBCO) global bathymetry dataset). 
Whales were less likely to move into waters less than 400m deep, and increasingly unlikely to move into even shallower waters. 
(iv) Phytoplankton concentration ($mmolNm^{-3}$, for combined diatom and non-diatom communities\cite{yool2013medusa}). 
This was included as a proxy for zooplankton food availability. 
Whales are more likely to move towards (or remain within) areas of high phytoplankton density, particularly during the foraging behavioural state.

We simulated the isotopic expression expected for (a) residency in each known hotspot for blue whale sightings or historic hunting grounds in the North Atlantic (Norwegian/Barents Sea, West Ireland, Canaries/Azores and Mid Atlantic Ridge\cite{mcdonald2006biogeographic,reilly2008balaenoptera,sigurjonsson1995life} ; Fig S4); (b) seasonal migration between high sub-Arctic latitudes and temperate latitudes around the British Isles and (c) seasonal migration between high latitudes and subtropical latitudes. 
For a full description of the model and its parameters see Supplementary Methods: Fig. S5 and Table S1. 
We simulated whale movements 50 times for each scenario and 30 times for each residency hotspot (Fig. S4). 
For each movement simulation we extracted phytoplankton $\delta^{13}$C values at each time point and location from the models described above\cite{magozzi2017using}. 
We then compared the simulated stable isotope profiles (Fig. \ref{fig3}) to the profile of the blue whale (Fig. \ref{fig1}).

\subsection{Data Availability}
Data are available from Zenodo doi: [DOI will be added on acceptance]. 
R code is available from GitHub (github.com/ALJackson/blue-whale-bes).
